### 2021-04-25 Chapter-1/Chapter-1.tex
(Introduction) Jim's edits

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
A side-channel attack is any attack that allows adversaries to infer sensitive information based on ``side-channel'' information. In 1996, Kocher published the first side-channel attack in the literature~\cite{kocher1996timing}. He demonstrated that an attacker could break cryptosystems like RSA or Diffie-Hellman by measuring the execution time of a simple modular exponentiation operation. The simple modular exponentiation calculates the remainder of an exponent divided by an integer.  The simple modular exponentiation operation executes faster when the exponent bit is 0 and slower when the exponent bit is 1. An attacker can infer the whole secret key by measuring the execution time of the operation. Since then, researchers have proposed numerous side-channel attacks which exploit numerous kinds of ``side-channel'' information, including execution timing~\cite{184415,disselkoen2017prime+}, power consumption~\cite{kar20178}, electromagnetic radiation~\cite{agrawal2002side,217605}, and even sound~\cite{chari1999towards}.

A large portion of side-channel attacks is the joint result of the underlying computer architecture design and the above software stack.  One example is cache channel attacks. The CPU cache is a smaller but faster memory designed to bridge the gap of the access time between the main memory and CPU registers by storing copies of recently accessed data. Because software may access different memory addresses when it processes different input data, and the size of the cache is significantly smaller than the size of the main memory, the CPU can only find part of the data in the cache. Such differences can be reflected in the differences between the program execution time. Hence we have cache side-channel attacks.

Various countermeasures have been proposed to defend against side-channel attacks. Based on the above discussion, existing countermeasures can be further divided into two categories, the hardware-based method and the software-based method. Hardware-level solutions, such as reducing shared resources, adopting oblivious RAM, and using transnational memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need to adopt new hardware features and make changes to today's modern complex computer systems, which are hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side channel vulnerabilities (data flow from secrets to load addresses, and data flow from secrets to branch conditions) from existing code base. For example, when we examine recent side-channel attacks, many of them exploit two kinds of code patterns: \emph{secret-dependent control-flow transfers} and \emph{secret-dependent data accesses}.
\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.32\textwidth, xrightmargin=.32\textwidth]
int secret[32];
...
while(i>0){
    r = (r * r) % n;
    if(secret[--i] == 1)
        r = (r * x) % n;   
}
\end{lstlisting}
    \caption{Secret-dependent control-flow transfers}
    \label{fig:secret:cf}
\end{figure}

Secret-dependent control-flow transfers happen when the value of secret can affect which branch the code executes. Figure~\ref{fig:secret:cf} shows an example. Depending on the value of \textsf{secret}, the code may or may not execute the code at line 6. As attackers, they can retrieve some information based on the execution control-flow.

@@

Secret-dependent data accesses happen when the value of the secret can affect which memory cell the code reads or writes. Figure~\ref{fig:secret:da} shows an example. \textsf{FSb} is an array. Depending on the value of \textsf{secret}, the code may read or write different items in the array.

The key intuition is that the above side-channel attacks happen when a program accesses different memory addresses if the program has different sensitive inputs. As shown in Figure~\ref{fig:secret:cf} and Figure~\ref{fig:secret:da}, if a program shows different patterns in control transfers or data accesses when the program processes different sensitive inputs, the program could have side-channel vulnerabilities. Various techniques can be used to exploit the above leakage site at different granularities. For example, cache channels can observe cache accesses at the level of cache sets~\cite{liu2015last}, cache lines~\cite{184415} or other granularities~\cite{yarom2017cachebleed}. Other types of side-channel attacks like controlled-channel attack~\cite{7163052}, can observe the memory access at the level of memory pages. So a natural defense approach is to remove the above code patterns in the software.

However, eliminating side-channel vulnerabilities from real-world software code base is not easy. First, side-channels are a low-level problem in computer systems. As a non-functional property, it can be difficult for programmers to reason about or test.  While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, they either suffer from expensive computation costs or high false positives and false negatives. Many of those reported vulnerabilities have not been fixed by developers for a variety of reasons. Second, most vulnerabilities pose a negligible risk. Although some vulnerabilities result in the key being compromised~\cite{184415, aumuller2002fault} entirely, many others are less severe in reality. Therefore, we need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.  Third, many side-channel vulnerable implementations (e.g., the CRT optimization and lookup tables) have better performance. It is a dilemma between the performance and the security.

@@

\begin{enumerate}
    \item It is hard to identify side-channel vulnerabilities in real-world software. In general, many side-channel vulnerabilities are low-level problems. Control-flows or memory accesses may not be side-channel vulnerabilities if they are independent of original secret inputs. Recent work can identify side-channel leakages. But those tools can only work on small programs or need to apply some domain knowledge of the target program to simplify the analysis.
    \item Side-channel vulnerabilities are ubiquitous. While some vulnerabilities can result in the key being entirely compromised, many other vulnerabilities are hard to exploit. Moreover, some vulnerable implementations perform better. Recent work can identify thousands of leaked points, but most of them are not patched by developers because those leakages are either false positives or leaked very little information.
\end{enumerate}

>>>>

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
A side-channel attack is any attack that allows adversaries to infer sensitive information based on ``side-channel'' information. In 1996, Kocher published the first side-channel attack in the literature~\cite{kocher1996timing}. He demonstrated that an attacker could break cryptosystems like RSA or Diffie-Hellman by measuring the execution time of a simple modular exponentiation operation. The simple modular exponentiation calculates the remainder of an exponent divided by an integer.  The simple modular exponentiation operation executes faster when the exponent bit is 0 and slower when the exponent bit is 1. An attacker can infer the whole secret key by measuring the execution time of the operation. Since then, researchers have proposed numerous side-channel attacks which exploit different kinds of ``side-channel'' information, including execution timing~\cite{184415,disselkoen2017prime+}, power consumption~\cite{kar20178}, electromagnetic radiation~\cite{agrawal2002side,217605}, and even sound~\cite{chari1999towards}.

A large portion of side-channel attacks is the joint result of the underlying computer architecture and the software stack.  One example is cache channel attacks. The CPU cache is a smaller but faster memory designed to bridge the gap of the access time between the main memory and CPU registers by storing copies of recently accessed data. Because software may access different memory addresses when it processes different input data, and the size of the cache is significantly smaller than the size of the main memory, the CPU only finds part of the data in the cache. Such differences can be reflected in the variance in the program execution time. Hence we have cache side-channel attacks.

Various countermeasures have been proposed to defend against side-channel attacks. Based on the above discussion, existing countermeasures can be further divided into two categories, the hardware-based method and the software-based method. Hardware-level solutions, such as reducing shared resources, oblivious RAM, and transnational memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need to adopt new hardware features and changes in today's modern complex computer systems, which are hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side-channel vulnerabilities (data flow from secrets to load addresses, and data flow from secrets to branch conditions) in existing code base. For example, when we examine recent side-channel attacks, many of them exploit two kinds of code patterns: \emph{secret-dependent control-flow transfers} and \emph{secret-dependent data accesses}.
\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.32\textwidth, xrightmargin=.32\textwidth]
int secret[32];
...
while(i>0){
    r = (r * r) % n;
    if(secret[--i] == 1)
        r = (r * x) % n;   
}
\end{lstlisting}
    \caption{Secret-dependent control-flow transfer}
    \label{fig:secret:cf}
\end{figure}

Secret-dependent control-flow transfers happen when the value of secret can affect which branch the code executes. Figure~\ref{fig:secret:cf} shows an example. Depending on the value of \textsf{secret}, the code may or may not execute line 6. An attacker can retrieve some information based on the execution control-flow.

@@

Secret-dependent data accesses happen when the value of the secret can affect which memory location the code reads or writes. Figure~\ref{fig:secret:da} shows an example. \textsf{FSb} is an array. Depending on the value of \textsf{secret}, the code may read or write different items in the array.

The key intuition is that the above side-channel attacks happen when a program accesses different memory addresses with different sensitive inputs. As shown in Figure~\ref{fig:secret:cf} and Figure~\ref{fig:secret:da}, if a program shows different patterns in control transfers or data accesses when the program processes different sensitive inputs, the program could have side-channel vulnerabilities. Various techniques can be used to exploit the above leakage site at different granularities. For example, cache channels can observe cache accesses at the level of cache sets~\cite{liu2015last}, cache lines~\cite{184415} or other granularities~\cite{yarom2017cachebleed}. Other types of side-channel attacks such as controlled-channel attack~\cite{7163052}, can observe a memory access at the level of memory pages. So a natural defense approach is to remove the above code patterns in the software.

However, eliminating side-channel vulnerabilities from real-world software code base is not easy. First, side-channels are a low-level problem in computer systems. As a non-functional property, it can be difficult for programmers to reason about or test.  While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, it either suffers from expensive computation costs or high false positives and false negatives.  Second, most vulnerabilities pose a negligible risk. Although some vulnerabilities result in the key being entirely compromised~\cite{184415, aumuller2002fault}, many others are less severe in reality. Therefore, we need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.  Third, many side-channel vulnerable implementations (e.g., the CRT optimization and lookup tables) have better performance. This is a dilemma between the performance and the security.

@@

\begin{enumerate}
    \item It is hard to identify side-channel vulnerabilities in real-world software. In general, many side-channel vulnerabilities are low-level problems. Control-flows or memory accesses may not be side-channel vulnerabilities if they are independent of the secret inputs. Recent work can identify side-channel leakages. But those tools can only work on small programs or need to apply domain knowledge of the target program to simplify the analysis.
    \item Side-channel vulnerabilities are ubiquitous. While some vulnerabilities can result in the key being entirely compromised, many other vulnerabilities are hard to exploit. Moreover, some vulnerable implementations perform better. Recent work can identify thousands of leaked points, but most of them are not patched by developers because those leakages are either false positives or leaked very little information.
\end{enumerate}
